The not so widening achievement gap: an interntational comparison of the evolution of the SES achievement gap from 2000-2015

### 2017-08-28 paper/paper.Rnw
Two models are complete - next writeup

# 2000: ST25Q01
# 1 <ISCED 2A>
# 2 <ISCED 2B>
# 3 <ISCED 2C>
# 4 <ISCED 3A>
# 5 <ISCED 3B>
# 6 <ISCED 3C>
# 7 N/A
# 8 M/R
# 9 Mis

# 2003: ISCEDO
# 2006: ISCEDO
# 2009: ISCEDO
# 2012: ISCEDO
# 2015: ISCEDO

# 1	General
# 2	Pre-Vocational
# 3	Vocational
# 4	Modular

library(readxl)

wd_extract <- function(var, name) {
  wdi_data <-
    WDI(indicator = var, start = 2000, end = 2015) %>%
    select(-1) %>%
    as_tibble() %>%
    mutate(year = as.character(year))

  names(wdi_data)[2] <- name
  wdi_data
}

pop_preprimary <- wd_extract("SP.PRE.TOTL.IN", "pop_preprimary")
teacher_pupil_ratio <- wd_extract("SE.PRE.ENRL.TC.ZS", "teacher_pupil_ratio")
perc_teacher_trained <- wd_extract("UIS.TRTP.0", "perc_teach")
preprimary_priv_enrol <- wd_extract("SE.PRE.PRIV.ZS", "preprimary_priv_enrol")

tracking_data <-
  readxl::read_xlsx("./paper/data/tracking.xlsx", sheet = "all_data") %>%
  map_if(is_double, round, 2) %>%
  as_tibble

@@

ready_data <-
  difference_data %>%
  left_join(tracking_data, by = c("country" = "cntry_name")) %>%
  left_join(central_examination) %>%
  mutate(num_tracks = ifelse(tracks15y == 1, 1, 0) %>% as.factor,
         age_selection = ifelse(selage >= 15, 1, 0)) %>%
  filter(!is.na(num_tracks), !is.na(age_selection), !is.na(difference))

@@

mod_tracking <-
  brms::brm(
  difference ~
    num_tracks +
    age_selection + (1 | country),
  family = gaussian(),
  data = ready_data,
  prior = c(set_prior("cauchy(2, 0.5)", coef = "num_tracks1"),
            set_prior("cauchy(13, 1)", coef = "age_selection")
            ),
  warmup = 1000, iter = 2000, chains = 5
  )

@@

new_data <-
  ready_data %>%
  filter(country %in% countries) %>%
  expand(country, age_selection, num_tracks) %>%
  filter(complete.cases(.))

@@

sum_ready_data <-
  ready_data %>%
  group_by(country, num_tracks, age_selection) %>%
  summarize(avg_diff = mean(difference),
            diff_low = avg_diff - 2 * sd(difference),
            diff_high = avg_diff + 2 * sd(difference)) %>%
  ungroup()

@@

pp_check(mod_tracking, nsamples = 100)

pp_check(mod_tracking, type = "error_scatter", nsamples = 100)

@@

mod_tracking <-
  ready_data_two %>%
  lm(cum_diff ~ num_tracks + age_selection, data = .) %>%
  arm::display()

mod_tracking <-
  ready_data_two %>%
  filter(!is.na(age_selection)) %>%
  brm(
    cum_diff ~ age_selection,
    family = gaussian(),
    data = .,
    prior = set_prior("cauchy(13, 1)", coef = "age_selection"),
    warmup = 1000, iter = 2000, chains = 5
  )

pp_check(mod_tracking, nsamples = 100)

>>>>

tracking_data <-
  read_xlsx("./paper/data/tracking.xlsx", sheet = "all_data") %>%
  map_if(is_double, round, 2) %>%
  as_tibble

@@

ready_data <-
  difference_data %>%
  left_join(tracking_data, by = c("country" = "cntry_name")) %>%
  left_join(central_examination) %>%
  left_join(school_data_ready) %>%
  mutate(num_tracks = ifelse(tracks15y == 1, 1, 0) %>% as.factor,
         age_selection = selage) %>%
  filter(!is.na(num_tracks), !is.na(age_selection), !is.na(difference))

ready_data_age <- mutate(ready_data, age_selection = ifelse(selage >= 15, 1, 0))

@@

mod_tracking <-
  brms::brm(
  difference ~
    num_tracks +
    age_selection + (1 | country),
  family = gaussian(),
  data = ready_data_age,
  prior = c(set_prior("cauchy(2, 0.5)", coef = "num_tracks1"),
            set_prior("cauchy(13, 1)", coef = "age_selection")
            ),
  warmup = 1000, iter = 2000, chains = 5
  )

@@

new_data <-
  ready_data_age %>%
  filter(country %in% countries) %>%
  expand(country, age_selection, num_tracks) %>%
  filter(complete.cases(.))

@@

sum_ready_data <-
  ready_data_age %>%
  group_by(country, num_tracks, age_selection) %>%
  summarize(avg_diff = mean(difference),
            diff_low = avg_diff - 2 * sd(difference),
            diff_high = avg_diff + 2 * sd(difference)) %>%
  ungroup()

@@

pp_check(mod_tracking, type = "error_binned", nsamples = 1000)
# The models tends to overpredict high values and underpredict low values.
@

@@

mod_tracking <-
  ready_data_two %>%
  lm(cum_diff ~ age_selection, data = .) %>%
  arm::display(detail = T)

mod_tracking <-
  ready_data_two %>%
  filter(!is.na(age_selection)) %>%
  brm(
    cum_diff ~ age_selection,
    family = gaussian(),
    data = .,
    prior = set_prior("cauchy(13, 1)", coef = "age_selection"),
    warmup = 1000, iter = 2000, chains = 5
  )
pp_check(mod_tracking, type = "dens_overlay", nsamples = 100)
pp_check(mod_tracking, type = "ecdf_overlay", nsamples = 10)
pp_check(mod_tracking, type = "boxplot", nsamples = 30)
pp_check(mod_tracking, type = "stat", nsamples = 200) # very nice to compare central tendency

pp_check(mod_tracking, type = "error_scatter_avg", nsamples = 1000)

pp_check(mod_tracking, type = "loo_pit")
# over prediction
# central tendency is not captured very well
# more under-prediction than over prediction
